Preparación de datos

### 2023-09-28 production/prod_prep.Rmd
re-estimacion con education distributive justice
- commit inicial desde fork frontiers
- nueva estimacion regresion
- solo promedio escolaridad como var nivel 2

load(url("https://dataverse.harvard.edu/api/access/datafile/6160173")) #ELSOC
save(elsoc_long_2016_2021,file = here::here("input/data-original/elsoc_long.Rdata"))
cit17<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216836"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit17,file = here::here("input/data-original/cit17.Rdata"))
cit18<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216941"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit18,file = here::here("input/data-original/cit18.Rdata"))
cit19<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216990"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit19,file = here::here("input/data-original/cit19.Rdata"))

@@

elsoc_long_2016_2021[elsoc_long_2016_2021 ==-999] <- NA
elsoc_long_2016_2021[elsoc_long_2016_2021 ==-888] <- NA
elsoc_long_2016_2021[elsoc_long_2016_2021 ==-777] <- NA
elsoc_long_2016_2021[elsoc_long_2016_2021 ==-666] <- NA

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2021 <- 
elsoc_long_2016_2021 %>% 
  #create new variables
  mutate(in_simpat=r09,
         sexo=m0_sexo,
         in_simili=r10,
         in_ansied=r11,
         in_valfam=r12_01,
         in_valami=r12_02,
         in_amsimb=r12_03,
         in_amreal=r12_04,
         in_acpier=r12_05,
         in_acadop=r12_06,
         in_amigcl=r12_07) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2021$in_simpat)
elsoc_long_2016_2021$in_simpat <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simpat, 
                      label = "Sympathy for [PER/VEN] living in Chile") 
elsoc_long_2016_2021$in_simpat <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simpat,
                       labels = c("Very little or not at all", "A little",
                                  "Somewhat","Quite a lot","A lot"))  
sjmisc::frq(elsoc_long_2016_2021$in_simpat)

#
sjmisc::frq(elsoc_long_2016_2021$in_simili)
elsoc_long_2016_2021$in_simili <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simili, 
                      label = "Similarity between Chileans and [PER/VEN]") 
elsoc_long_2016_2021$in_simili <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simili,
                       labels = c("Not similar", "Slightly similar", "Somewhat similar",
                                  "Quite similar", "Very similar"))
sjmisc::frq(elsoc_long_2016_2021$in_simili)

#
sjmisc::frq(elsoc_long_2016_2021$in_ansied)
elsoc_long_2016_2021$in_ansied <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_ansied, 
                      label = "Interaction anxiety with [PER/VEN].") 
elsoc_long_2016_2021$in_ansied <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_ansied,
                       labels = c("Very uncomfortable", "Uncomfortable", 
                                  "Neither comfortable nor uncomfortable",
                                  "Comfortable", "Very comfortable"))
sjmisc::frq(elsoc_long_2016_2021$in_ansied)

for (i in c("in_valfam","in_valami","in_amsimb","in_amreal",
            "in_acpier","in_acadop","in_amigcl")) {
elsoc_long_2016_2021[[i]] <-
sjlabelled::set_labels(x = elsoc_long_2016_2021[[i]],
                       labels = c("Strongly disagree","Disagree",
                                  "Neither disagree nor agree",
                                  "Agree", "Strongly agree"))
}

sjmisc::frq(elsoc_long_2016_2021$in_valfam)
elsoc_long_2016_2021$in_valfam <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valfam, 
                      label = "Family value of migrant friends [PER/VEN]") 
sjmisc::frq(elsoc_long_2016_2021$in_valfam)

sjmisc::frq(elsoc_long_2016_2021$in_valami)
elsoc_long_2016_2021$in_valami <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valami, 
                      label = "Friends value of migrant friends [PER/VEN]") 
sjmisc::frq(elsoc_long_2016_2021$in_valami)

sjmisc::frq(elsoc_long_2016_2021$in_amsimb)
elsoc_long_2016_2021$in_amsimb <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amsimb, 
                      label = "Loose of identity because of migrants (reverse)")
sjmisc::frq(elsoc_long_2016_2021$in_amsimb) 
# Revert item
elsoc_long_2016_2021$in_amsimb<- 
  sjmisc::rec(elsoc_long_2016_2021$in_amsimb,rec = "rev")
sjmisc::frq(elsoc_long_2016_2021$in_amsimb) # check reverted item

sjmisc::frq(elsoc_long_2016_2021$in_amreal)
elsoc_long_2016_2021$in_amreal <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amreal, 
                      label = "Unemployment increases") 
sjmisc::frq(elsoc_long_2016_2021$in_amreal)

sjmisc::frq(elsoc_long_2016_2021$in_acpier)
elsoc_long_2016_2021$in_acpier <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acpier, 
                      label = "Migrant keeping their customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acpier)

sjmisc::frq(elsoc_long_2016_2021$in_acadop)
elsoc_long_2016_2021$in_acadop <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acadop, 
                      label = "Adoption of Chilean customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acadop)

sjmisc::frq(elsoc_long_2016_2021$in_amigcl)
elsoc_long_2016_2021$in_amigcl <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amigcl, 
                      label = "migrants [PER/VEN] have Chilean friends") 
sjmisc::frq(elsoc_long_2016_2021$in_amigcl)

# Average variables___________________________________________________________

elsoc_long_2016_2021 <- elsoc_long_2016_2021 %>% 
  mutate(conviviality=as.numeric((in_ansied+in_simpat+in_valfam+in_valami+in_amigcl)/5),
         identity= as.numeric((in_simili+in_amsimb+in_acpier+in_acadop)/4),
         threat = as.numeric(in_amreal))

sjmisc::frq(elsoc_long_2016_2021$conviviality)
elsoc_long_2016_2021$conviviality <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$conviviality, 
                      label = "Conviviality")
sjmisc::frq(elsoc_long_2016_2021$conviviality)

sjmisc::frq(elsoc_long_2016_2021$identity)
elsoc_long_2016_2021$identity <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$identity, 
                      label = "Identity")
sjmisc::frq(elsoc_long_2016_2021$identity)

# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2021$educ <- 
  car::recode(elsoc_long_2016_2021$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2021$educ <-
  factor(elsoc_long_2016_2021$educ,
         labels = c("Primary","High school","Technical","Universitary"))

#reverse education, reference level is the highest level
elsoc_long_2016_2021$educ <- forcats::fct_rev(elsoc_long_2016_2021$educ)

elsoc_long_2016_2021$educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$educ,
                      label = "Education")
sjmisc::frq(elsoc_long_2016_2021$educ)

#Recoding of education to years based on casen 2017.
elsoc_long_2016_2021$educyear<- as.numeric(
  car::recode(elsoc_long_2016_2021$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

elsoc_long_2016_2021$educyear <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$educyear,
                      label = "Education in years")

class(elsoc_long_2016_2021$educyear)
sjmisc::frq(elsoc_long_2016_2021$educyear)

@@

#Impute midpoint of income ranges
elsoc_long_2016_2021$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2021$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2021$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2021$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2021$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2021$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2021$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2021$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2021$m29_rec),
         yes =  elsoc_long_2016_2021$m29_rec,
         no =  elsoc_long_2016_2021$m30_rec)
summary(elsoc_long_2016_2021$m29_imp)

elsoc_long_2016_2021$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2021$m29_imp),
         yes =  elsoc_long_2016_2021$m30b_rec,
         no =  elsoc_long_2016_2021$m29_imp)
summary(elsoc_long_2016_2021$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2021$deflactor <-
  with(elsoc_long_2016_2021, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2021 <-
  elsoc_long_2016_2021 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54))
sjmisc::frq(elsoc_long_2016_2021$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2021$n_hogar_r<-
  car::recode(elsoc_long_2016_2021$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2021$ing_pc <- 
  (elsoc_long_2016_2021$m29_imp/elsoc_long_2016_2021$n_hogar_r)

elsoc_long_2016_2021$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2021$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2021$ing_pc)

# Compute income quintiles
elsoc_long_2016_2021 <- elsoc_long_2016_2021 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2021$quintil <- 
  factor(elsoc_long_2016_2021$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
elsoc_long_2016_2021$quintil <- forcats::fct_rev(elsoc_long_2016_2021$quintil)

elsoc_long_2016_2021$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2021$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2021$quintil1<-
  car::recode(elsoc_long_2016_2021$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

elsoc_long_2016_2021$quintil1 <- factor(elsoc_long_2016_2021$quintil1, c("Q5","Q4","Q3","Q2","Q1","QNA"))

elsoc_long_2016_2021$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2021$quintil1)

#Age Ranges
elsoc_long_2016_2021$edad <- 
  factor(car::recode(elsoc_long_2016_2021$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2021$edad <-
  sjlabelled::set_label(elsoc_long_2016_2021$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2021$ess <- 
as.numeric(elsoc_long_2016_2021$d01_01)
sjmisc::frq(elsoc_long_2016_2021$ess)
elsoc_long_2016_2021$ess <- sjmisc::rec(elsoc_long_2016_2021$ess,rec = "rev")
sjmisc::frq(elsoc_long_2016_2021$ess)

elsoc_long_2016_2021$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$ess,
                      label = "Subjective Social Status: individual (reverse)")

sjmisc::frq(elsoc_long_2016_2021$ess)
# sjPlot::plot_frq(elsoc_long_2016_2021$ess)
# Subjective social status: migrants (w04 and w05)____________________________
elsoc_long_2016_2021$ess_inm <- 
as.numeric(elsoc_long_2016_2021$d01_04)

sjmisc::frq(elsoc_long_2016_2021$ess_inm)
elsoc_long_2016_2021$ess_inm <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$ess_inm,
                      label = "Subjective Social Status: inmigrants")

sjmisc::frq(elsoc_long_2016_2021$ess_inm)
# sjPlot::plot_frq(elsoc_long_2016_2021$ess_inm)

# Nationality of the interviewee____________________________

elsoc_long_2016_2021$nation <- 
elsoc_long_2016_2021$m45
sjmisc::frq(elsoc_long_2016_2021$nation)

elsoc_long_2016_2021$nation <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$nation, 
                      label = "Nationality") 
elsoc_long_2016_2021$nation <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$nation,
                       labels = c("Chilean","Peruvian","Argentina","Colombian",
                                  "Bolivian", "Haitian","Spanish","Other"))

sjmisc::frq(elsoc_long_2016_2021$nation)

# Known: Immigrant [PER/VEN]____________________________

# sjmisc::find_var(elsoc_long_2016_2021,"Conocidos")
sjmisc::frq(elsoc_long_2016_2021$r05_01)
elsoc_long_2016_2021$know_inm <- 
elsoc_long_2016_2021$r05_01

elsoc_long_2016_2021$know_inm <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$know_inm, 
                      label = "Known: Immigrants") 
elsoc_long_2016_2021$know_inm <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$know_inm,
                       labels = c("None","Few","Some","Quite a few","Many"))
sjmisc::frq(elsoc_long_2016_2021$know_inm)

# Level of contact: Migrant friends_________________________
sjmisc::frq(elsoc_long_2016_2021$r05_02)
elsoc_long_2016_2021$frien_inm <- 
elsoc_long_2016_2021$r05_02

elsoc_long_2016_2021$frien_inm <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$frien_inm, 
                      label = "Level of contact: Inmigrant friends")

elsoc_long_2016_2021$frien_inm <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$frien_inm,
                       labels = c("None","Few","Some","Quite a few","Many"))
sjmisc::frq(elsoc_long_2016_2021$frien_inm)

# Political position ____________________________
elsoc_long_2016_2021$pos_id <-
factor(
  car::recode(
    elsoc_long_2016_2021$c15,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

elsoc_long_2016_2021$pos_id <- factor(elsoc_long_2016_2021$pos_id,
                                      levels = levels(elsoc_long_2016_2021$pos_id))

elsoc_long_2016_2021$pos_id <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$pos_id, 
                      label = "Political identification")

sjmisc::frq(elsoc_long_2016_2021$pos_id)

# CONTACT_______________________________________________________________

# NATION to wave 1 (original)___________________
df_nation <- elsoc_long_2016_2021 %>%  
  select(idencuesta,nation_fix=nation,ola,muestra) %>% 
  filter(ola %in% c(1,3)) %>% #keep wave 1 and 3 
  group_by(idencuesta) %>% 
  summarise(nation_fix=max(nation_fix,na.rm = F)) #keep the maximum value 
sjmisc::frq(df_nation$nation_fix) # 2 NA

elsoc <- elsoc_long_2016_2021 #dummy dataset _________________________________

# merge the data with fixed variable:
elsoc <- left_join(x = elsoc,y = df_nation,by ="idencuesta") 
# sjmisc::frq(elsoc$nation_fix)

table(elsoc$nation==elsoc$nation_fix) #how many cases are different ?
elsoc$check_nation <- ifelse(elsoc$nation==elsoc$nation_fix,
                             yes = "IGUALES",
                             no = "DISTINTOS")
# sjmisc::frq(elsoc$check_nation) #18 cases

# edit, if value is different from original variable, set the value of wave 1
elsoc$nation_t1 <- if_else(condition = elsoc$check_nation=="DISTINTOS",
                           true = elsoc$nation,
                           false = elsoc$nation_fix,
                           missing = elsoc$nation_fix)

table(elsoc$nation==elsoc$nation_t1)
sjmisc::frq(elsoc$nation_t1)

>>>>

load(url("https://dataverse.harvard.edu/api/access/datafile/7245118")) #ELSOC
save(elsoc_long_2016_2022.2,file = here::here("input/data-original/elsoc_long.Rdata"))

cit17<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216836"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit17,file = here::here("input/data-original/cit17.Rdata"))
cit18<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216941"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit18,file = here::here("input/data-original/cit18.Rdata"))
cit19<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216990"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit19,file = here::here("input/data-original/cit19.Rdata"))

@@

elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-999] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-888] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-777] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-666] <- NA

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2022.2 <- 
elsoc_long_2016_2022.2 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
elsoc_long_2016_2022.2$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_educ, 
                      label = "Education distributive justice")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
elsoc_long_2016_2022.2$merit_effort <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_effort, 
                      label = "People are rewarded for their efforts")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
elsoc_long_2016_2022.2$merit_talent <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_talent, 
                      label = "People are rewarded for their intelligence")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
elsoc_long_2016_2022.2$ahead_family <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_family, 
                      label = "Coming from a wealthy family")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
elsoc_long_2016_2022.2$ahead_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_educ, 
                      label = "Having a good level of education")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
elsoc_long_2016_2022.2$ahead_ambition <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_ambition, 
                      label = "Having ambition")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
elsoc_long_2016_2022.2$ahead_hardwork <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_hardwork, 
                      label = "Hard work")

for (i in c("just_educ","merit_effort","merit_talent",
            "ahead_family","ahead_educ","ahead_ambition", "ahead_hardwork")) {
elsoc_long_2016_2022.2[[i]] <-
sjlabelled::set_labels(x = elsoc_long_2016_2022.2[[i]],
                       labels = c("Strongly disagree","Disagree",
                                  "Neither disagree nor agree",
                                  "Agree", "Strongly agree"))
}

sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2022.2$educ <- 
  car::recode(elsoc_long_2016_2022.2$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2022.2$educ <-
  factor(elsoc_long_2016_2022.2$educ,
         labels = c("Primary","High school","Technical","Universitary"))

#reverse education, reference level is the highest level
#elsoc_long_2016_2022.2$educ <- forcats::fct_rev(elsoc_long_2016_2022.2$educ)

elsoc_long_2016_2022.2$educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educ,
                      label = "Education")
sjmisc::frq(elsoc_long_2016_2022.2$educ)

#Recoding of education to years based on casen 2017.
elsoc_long_2016_2022.2$educyear<- as.numeric(
  car::recode(elsoc_long_2016_2022.2$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

elsoc_long_2016_2022.2$educyear <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educyear,
                      label = "Education in years")

class(elsoc_long_2016_2022.2$educyear)
sjmisc::frq(elsoc_long_2016_2022.2$educyear)

@@

#Impute midpoint of income ranges
elsoc_long_2016_2022.2$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2022.2$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2022.2$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2022.2$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2022.2$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2022.2$m29_rec),
         yes =  elsoc_long_2016_2022.2$m29_rec,
         no =  elsoc_long_2016_2022.2$m30_rec)
summary(elsoc_long_2016_2022.2$m29_imp)

elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2022.2$m29_imp),
         yes =  elsoc_long_2016_2022.2$m30b_rec,
         no =  elsoc_long_2016_2022.2$m29_imp)
summary(elsoc_long_2016_2022.2$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2022.2$deflactor <-
  with(elsoc_long_2016_2022.2, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2022.2 <-
  elsoc_long_2016_2022.2 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54))
sjmisc::frq(elsoc_long_2016_2022.2$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2022.2$n_hogar_r<-
  car::recode(elsoc_long_2016_2022.2$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2022.2$ing_pc <- 
  (elsoc_long_2016_2022.2$m29_imp/elsoc_long_2016_2022.2$n_hogar_r)

elsoc_long_2016_2022.2$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2022.2$ing_pc)

# Compute income quintiles
elsoc_long_2016_2022.2 <- elsoc_long_2016_2022.2 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2022.2$quintil <- 
  factor(elsoc_long_2016_2022.2$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
#elsoc_long_2016_2022.2$quintil <- forcats::fct_rev(elsoc_long_2016_2022.2$quintil)

elsoc_long_2016_2022.2$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2022.2$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2022.2$quintil1<-
  car::recode(elsoc_long_2016_2022.2$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#elsoc_long_2016_2022.2$quintil1 <- factor(elsoc_long_2016_2022.2$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

elsoc_long_2016_2022.2$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2022.2$quintil1)

#Age Ranges
elsoc_long_2016_2022.2$edad <- 
  factor(car::recode(elsoc_long_2016_2022.2$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2022.2$edad <-
  sjlabelled::set_label(elsoc_long_2016_2022.2$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2022.2$ess <- 
as.numeric(elsoc_long_2016_2022.2$d01_01)
sjmisc::frq(elsoc_long_2016_2022.2$ess)
#elsoc_long_2016_2022.2$ess <- sjmisc::rec(elsoc_long_2016_2022.2$ess,rec = "rev")
#sjmisc::frq(elsoc_long_2016_2022.2$ess)

elsoc_long_2016_2022.2$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ess,
                      label = "Subjective Social Status: individual")

sjmisc::frq(elsoc_long_2016_2022.2$ess)
# sjPlot::plot_frq(elsoc_long_2016_2022.2$ess)

# Political position ____________________________
elsoc_long_2016_2022.2$pos_id <-
factor(
  car::recode(
    elsoc_long_2016_2022.2$c15,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

elsoc_long_2016_2022.2$pos_id <- factor(elsoc_long_2016_2022.2$pos_id,
                                      levels = levels(elsoc_long_2016_2022.2$pos_id))

elsoc_long_2016_2022.2$pos_id <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$pos_id, 
                      label = "Political identification")

sjmisc::frq(elsoc_long_2016_2022.2$pos_id)

# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2022.2 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

# Reshape long to wide
df_study1_wide <- df_study1 %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1,comuna,comunacod,just_educ:edad,sexo))
                     )

# fix data to w01 values
df_study1_wide$edad_2 <-df_study1_wide$edad_1 #age
df_study1_wide$edad_3 <-df_study1_wide$edad_1
df_study1_wide$edad_4 <-df_study1_wide$edad_1
df_study1_wide$edad_5 <-df_study1_wide$edad_1
df_study1_wide$edad_6 <-df_study1_wide$edad_1

df_study1_wide$sexo_2 <-df_study1_wide$sexo_1 #sex
df_study1_wide$sexo_3 <-df_study1_wide$sexo_1
df_study1_wide$sexo_4 <-df_study1_wide$sexo_1
df_study1_wide$sexo_5 <-df_study1_wide$sexo_1
df_study1_wide$sexo_6 <-df_study1_wide$sexo_1

df_study1_wide$educ_2 <-df_study1_wide$educ_1 #education
df_study1_wide$educ_3 <-df_study1_wide$educ_1
df_study1_wide$educ_4 <-df_study1_wide$educ_1
df_study1_wide$educ_5 <-df_study1_wide$educ_1
df_study1_wide$educ_6 <-df_study1_wide$educ_1

df_study1_wide$educyear_2 <-df_study1_wide$educyear_1 #education years
df_study1_wide$educyear_3 <-df_study1_wide$educyear_1
df_study1_wide$educyear_4 <-df_study1_wide$educyear_1
df_study1_wide$educyear_5 <-df_study1_wide$educyear_1
df_study1_wide$educyear_6 <-df_study1_wide$educyear_1

df_study1_wide$quintil1_2 <-df_study1_wide$quintil1_1 #quintiles
df_study1_wide$quintil1_3 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_4 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_5 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_6 <-df_study1_wide$quintil1_1

df_study1_wide$ess_2 <-df_study1_wide$ess_1 # subjective status
df_study1_wide$ess_3 <-df_study1_wide$ess_1
df_study1_wide$ess_4 <-df_study1_wide$ess_1
df_study1_wide$ess_5 <-df_study1_wide$ess_1
df_study1_wide$ess_6 <-df_study1_wide$ess_1

df_study1_wide$pos_id_2 <-df_study1_wide$pos_id_1 # political position
df_study1_wide$pos_id_3 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_4 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_5 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_6 <-df_study1_wide$pos_id_1
